---
title: "HW6_Lab11_sample_code"
author: "Qian Xiao"
date: '2020-04-13'
output: html_document
---
#Packages 
```{r}
library(FirebrowseR) 
library(bladderbatch)
library(limma)
library(sva)
library(dplyr)
library(survival)
library(glmnet)
library(ggplot2)
library(survminer) 
```

#PartI Q2
```{r}
# download all available cohorts
cohorts <- Metadata.Cohorts(format = "csv")
# show what cohorts are available
head(cohorts)
```

```{r}
#Retrieve a list of 150 patients associated with a cancer type
head(Samples.Clinical(cohort="BRCA",format="tsv"))
```


```{r}
# Download BRCA data (have to do this because we can only receive 150 patients at a time)
all.Received <- FALSE
page.Counter <- 1
page.size <- 150
brca_pats <- list()
while(all.Received == FALSE) {
    brca_pats[[page.Counter]] <- Samples.Clinical(format = "csv",
            cohort = "BRCA", page_size = page.size, page = page.Counter)
    if(page.Counter > 1) {
        colnames(brca_pats[[page.Counter]]) <-
            colnames(brca_pats[[page.Counter-1]])
    }

    if(nrow(brca_pats[[page.Counter]]) < page.size) {
        all.Received = TRUE
    } else {
        page.Counter = page.Counter + 1
    }
}

brca_pats <- do.call(rbind, brca_pats)

#Dimension of the table containing patient info
dim(brca_pats)
```

```{r}
head(brca_pats)
```


#PartII Hints

##LIMMA
```{r, eval=FALSE}

lmFit() #For fitting LIMMA

eBayes() #For getting summary statistics

topTable() #Wrapping summary statistics up into a table, adjusted P is equivalent to FDR, should consider the absolute value of logFC because we want both UPREGULATED and DOWNREGULATED genes

```

##Survival Analysis

```{r}
# using lung data from survival package
# data wrangling to make this easier
lung2 <- lung
# 1 = died, 0 = still alive at last observation
lung2$death <- lung$status - 1
lung2$sex <- lung2$sex - 1

km_fit <- survfit(Surv(time, death) ~ sex, data = lung2)
ggsurvplot(km_fit, pval = TRUE, pval.method = TRUE) +
    ggtitle("Survival for Sexes")
```

##Cox Regression
```{r}
cox_mod1 <- coxph(Surv(time, death) ~ sex, data = lung2)
summary(cox_mod1)
```

##LASSO regularization
```{r}
set.seed(321)

lung_nona <- na.omit(lung2) #Omit NAs

x <- as.matrix(lung_nona[,4:10]) #Subset the covariates and assign it to x

survobj <- Surv(lung_nona$time, lung_nona$death) #Create the Survival object, which is the outcome we are interested in

cvfit <- cv.glmnet(x, survobj, family = "cox", alpha = 1) #Cross-validation

cvfit
```

```{r}
coef(cvfit, s = "lambda.min") #The set of coefficients using lambda.min, choose those that are non-zero to run cox regression
```

##Random cox models VS differential gene cox model

The idea is to have a for loop. For each iteration, randomly sample n predictors from all the predictors, subset the data with outcome and the selected n predictors, and then run cox regression using the subsetted data frame. You may record AIC/BIC of each random cox model for downstream comparison.



