---
title: "Lab 5 Relevant Code"
author: "Jiazhen Rong, Qing Xiao"
date: '2021-02-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

### Please notice that the following codes are for Lab 5 session. Please look at the code with Lab 5 recorded viedos and slides for better understanding. For live demos of using the webpage-based tools, please view Lab 5 recorded viedos.
* This Lab covers Part I-IV of HW3. 
* Some codes were adapted from previous year's YA Qing Xiao.

## Bash Script Review

```{r, engine='bash', eval=FALSE}
#!/bin/bash
#SBATCH -n 1 # Number of cores requested
#SBATCH -N 1 # Ensure that all cores are on one machine
#SBATCH -t 30 # Runtime in minutes
#SBATCH -p serial_requeue # Partition to submit to
#SBATCH --mem=32G # Memory in GB (see also --mem-per-cpu)
#SBATCH -o p1.out # Standard out goes to this file
#SBATCH -e p1.err # Standard err goes to this file

# LOAD_MODULES
module load <module_name>

Your code….

```

## How to Insert Figures in .Rmd
```{r, engine='bash', eval=FALSE}
{r,echo=FALSE,out.width = '100%'}  (add ``` at the front of the line)
knitr::include_graphics("/path/to/your/png")
```

## BWA Alignment and Samtools for Summary Statistics: 

```{r, engine='bash', eval=FALSE}
# your shebang/bash script
# Check the version to use by using `module spider <module's name>` or `module avail`
module load bwa/0.7.15-fasrc02
bwa mem /path/to/index/fasta /path/to/input/data > /path/to/output/file/your_output_name.sam
```

```{r, engine='bash', eval=FALSE}
# samtools might be useful to acquire the summary statistics
module load samtools/1.5-fasrc02

#Check the number of total reads and successfully mapped reads
$ samtools flagstat bwa.sam 
#Create a bam file of uniquely mapped reads
$ samtools view -bq 1 bwa.sam > unique.bam
#Again check the unique bam file to find the number of uniquelly mapped reads
$ samtools flagstat unique.bam

#Then calculate proportion of uniquely mapped from the output file
```

## MACS Remove Duplicated Reads:

```{r, engine='bash', eval=FALSE}

# your shebang
module load centos6/0.0.1-fasrc01
module load macs2/2.1.2_dev-fasrc01

macs2 filterdup -i /path/to/input/bam/file -g hs --keep-dup 1 -o ./path/to/output/bed/file/your_output_name.bed

#You may find the % redundancy in the .err file
```

## MACS2 - Peak Calling for one sample:
Call peak for one sampleL

```{r, engine='bash', eval=FALSE}
# your shebang/bash script
module load centos6/0.0.1-fasrc01
module load macs2/2.1.2_dev-fasrc01

macs2 callpeak -t /path/to/your/input/sample/bed/file.bed -f AUTO -g hs -q <FDR cutoff> --fe-cutoff <fold change> --outdir path/to/save/your/output/ -n prefix_of_your_output

#The fragment length can also be found in .err files
#then Use `wc -l` to count the number of peaks
```

```{r, engine='bash', eval=FALSE}
-t/--treatment filename
-c/--control
-n/--output name
-f/--format of tag files
--outdir/--the folder where all the output files saved into
-n/--name of the output as NAME_peaks.bed
-g/--gsize The default hs -- 2.7e9 is recommended as_for UCSC human hg18 assembly
-q/--qvalue (minimum FDR) cutoff to call significant regions. Default is 0.05.
```

Run the above commands for both tumor and normal sample.

## MACS2 - Peak Calling with control (using normal sample as control):

```{r, engine='bash', eval=FALSE}
# your shebang
module load centos6/0.0.1-fasrc01
module load macs2/2.1.2_dev-fasrc01

macs2 callpeak -t path/to/your/treat.bed -c path/to/your/control.bed -f AUTO -g hs -q <FDR cutoff> --fe-cutoff <fold change> --outdir path/to/your/output/folder/ -n prefix_of_your_output
```
The output of prefix_of_your_output_summits.bed will be used for downstream analysis.


## Rules for submitting the homework:

Please submit your solution directly on the canvas website. Please
provide both your code in this Rmd document and an html file for your
final write-up. Please pay attention to the clarity and cleanness of
your homework.

The teaching fellows will grade your homework and give the grades with
feedback through canvas within one week after the due date. Some of the
questions might not have a unique or optimal solution. TFs will grade
those according to your creativity and effort on exploration, especially
in the graduate-level questions.

## Acknowledgement
- Dr. X. Shirley Liu
- Qing Xiao
