---
title: "2021 Lab 6 Relevant Code"
author: "Jiazhen Rong, Qing Xiao"
date: '2021-03-02'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

### Please notice that the following codes are for Lab 6 session. Please look at the code with Lab 6 recorded viedos and slides for better understanding. For live demos of using the webpage-based tools, please view Lab 6 recorded viedos.
* This Lab covers Part V-VI of HW3. 
* Some codes were adapted from previous year's TA Qing Xiao.

## Re-write RefGene Annotation to BED format with Python:

```{python,eval=FALSE}
import pandas as pd

#Read in the reference data
TSS = pd.read_table('/path/to/refGene.txt', header=None)

#Create a subset from it with the columns we need 
TSS_sub = TSS.iloc[:,[2,4,5,3,12]]
TSS_sub.columns = ["chr","start","end","strand","id"]

#Create subsets of genes on positive strand and negative strand
TSS_pos = TSS_sub[TSS_sub.strand == '+']
TSS_neg = TSS_sub[TSS_sub.strand == '-']

#Rearrange the data and create a new dataframe containing the information of TSS we want
TSS_pos.end = TSS_pos.start
TSS_neg.start = TSS_neg.end
TSS_new = TSS_pos.append(TSS_neg)
TSS_new.end = TSS_new.end + 1

#Export it into a bed file
TSS_new.to_csv("/your/TSS_hg38_bed_file",sep='\t',header=False, index=False)
```

Then you can use BEDTools window for finding which hg38 genes have ChIP-Seq peaks within +/- 2KB windows of their TSS.

```{r, engine='bash', eval=FALSE}
module load centos6/0.0.1-fasrc01
module load bedtools/2.17.0-fasrc01

#We can use bedtools to select a window size, then find the overlaps between the peaks and the annotated genes
bedtools window -w <window_size> -u -a <your_new_tss_bed_file> -b q4_diff_peaks_summits.bed > <your_output_bed>

#Count the number of genes
wc -l <your_output_bed>

#The same task might be also completed with the python package wrapping BEDTools (pybedtools), but you would need to install and import it 
```


If you want to submit the script on server instead of running it on your local computer, you can wrap the python code in main function, and save it in a .py script. Then you can submit the python script with bash script.
```{r,eval=FALSE}
#### Within you q12.py script

def main():
        #Read in the reference data
        TSS = pd.read_table('/path/to/refGene.txt', header=None)

        #Create a subset from it with the columns we need 
        TSS_sub = TSS.iloc[:,[2,4,5,3,12]]
        TSS_sub.columns = ["chr","start","end","strand","id"]

        #Create subsets of genes on positive strand and negative strand
        TSS_pos = TSS_sub[TSS_sub.strand == '+']
        TSS_neg = TSS_sub[TSS_sub.strand == '-']

        #Rearrange the data and create a new dataframe containing the information of TSS we want
        TSS_pos.end = TSS_pos.start
        TSS_neg.start = TSS_neg.end
        TSS_new = TSS_pos.append(TSS_neg)
        TSS_new.end = TSS_new.end + 1

        #Export it into a bed file
        TSS_new.to_csv("/your/TSS_hg38_bed_file",sep='\t',header=False, index=False)

if __name__ == '__main__':
    main()
```

```{r,eval=FALSE}
# LOAD_MODULES
module load python/3.7.7-fasrc01
python q12.py
```

## For getting Gene Names to run DAVID:
In Q13 & 14, you will need to extract gene names (some specific columns) out of some files.
Sometimes you will need to rank each gene by scores, finding overlaps with other gene lists and in the end you need to output gene names for DAVID.
* Please use and select Official Gene Symbol names when running DAVID.

```{r, eval=FALSE}
# check the following commands in R:
?read.table
?order
?intersect
?write.table
```

For Q13:
```{r, eval=FALSE}
#Read in your Q12 result and extract official gene symbols

#Run DAVID with these putative target genes alone

#Find the overlap between the putative target genes and the upregulated genes in prostate cancer provided in Q13

#Run DAVID again with these overlapping genes
```

For Q14:
```{r, eval=FALSE}
#Read in the peaks with different regulatory potential score

#Rank them by the score and select the top 1500 genes, then run DAVID

#Find the overlap between the top 1500 genes and the upregulated genes you imported in Q13

#Generate lists of genes and run DAVID again
```


## ATAC-seq
Getting Differential ATAC-Seq peaks with BEDTools: <br/>
* Take a look at bedtools subtract <br/>
* please include -A option <br/>
* You need to cut the first 3 columns otherwise downstream process might complain <br/>

```{r, engine='bash', eval=FALSE}
module load centos6/0.0.1-fasrc01
module load bedtools/2.17.0-fasrc01

your bedtool command | cut -f 1-3 > <your_out_put.bed>
```

## Running LISA on server:

```{r,eval=FALSE}
# load module
module load Anaconda/5.0.1-fasrc02

# use pre-installed lisa2 & acticate environment
source /n/stat115/2021/HW5/miniconda3/bin/activate
conda activate lisa2

lisa oneshot hg38 <your/up_regulated_gene_names_file/txt> -b 300 --seed=2556 --save_metadata > </your/output/txt>
```
Then take a screen shot of the top TFs that might be regulators given the upregualted gene list.

## Rules for submitting the homework:

Please submit your solution directly on the canvas website. Please
provide both your code in this Rmd document and an html file for your
final write-up. Please pay attention to the clarity and cleanness of
your homework.

The teaching fellows will grade your homework and give the grades with
feedback through canvas within 1-2 week after the due date. Some of the
questions might not have a unique or optimal solution. TFs will grade
those according to your creativity and effort on exploration, especially
in the graduate-level questions.

## Acknowledgement
- Dr. X. Shirley Liu
- Qing Xiao



